Investigation of the activities performed by experimental researchers in a software engineering lab using an ethnographic methodology

### 2024-02-18 revisions/v3/Response-to-the-reviewers-comments-2.md

# Responses to the Reviewers' comments

We appreciate the editor and reviewers for the comments we have received. We have considered all the comments in the current revision. The improvements are indicated in the manuscript using track changes.

The final version of the manuscript can be obtained by replacing the line:

> \usepackage[draft]{changes}

 in the "Article.tex" file with:

> \usepackage[final]{changes}

The manuscript is hosted on the following GitHub repository:

> https://github.com/rodrifonsi/ESE-Process

In some cases, the changes could not be highlighted in the document due to limitations of the *changes* package. However, these changes can be identified by performing a diff between the HEAD and HEAD~1 versions of the .tex files using the following command:

> git diff

Below, we detail the responses to the reviewers' comments and the changes made:

**Comment 1** - Please add recent references.

*Response to Comment 1* - We have included the following recent references:

> Shepperd, M., Ajienka, N., Counsell, S.: ‘The role and value of replication in empirical software engineering results’, Information and Software Technology, 2018, 99, pp. 120–132

> Cockburn, A., Dragicevic, P., Besançon, L., Gutwin, C.: ‘Threats of a replication crisis in empirical computer science’, Communications of the ACM, 2020, 63, (8), pp. 70–79

> The International Organization for Standardization: ‘ISO/IEC 20000-1:2018 Information Technology - Service Management - Part 1: Service Management System Requirements’. 2018. https://www.iso.org/standard/70636.html

>European Commission. ‘Researchcomp: The european competence framework for researchers’. 2022. Accessed: 2024-02-18. https://research-and-innovation.ec.europa.eu/document/download/7da29338-37bf-4d51-b5eb-a1571b84c7ad_en?filename=ec_rtd_research-competence-presentation.pdf

**Comment 2** - Add some future scope conclusion section.

*Response to Comment 2* - Thank you for your suggestion. We have created a Section titled *Implications for practice* containing the following information:

> Our findings suggest specific improvements that could be applied to SE experimentation. The first and most evident is transitioning from an experimental process primarily based on the application of experimental designs and the conduct of statistical analyses to a more professional approach based on explicit processes, activities, and roles, as is customary in many areas, e.g., Information Technology \cite{iso20000}.
>
>It could be argued that current SE experimentation already has a well-defined process and activities, as can be seen, e.g., in the works of Juristo \& Moreno \cite{Juristo-2001-SE-experimentation} or Wohlin \textit{et al.} \cite{Wohlin-2000-Experimentation-SE}. However, as we have argued before, such a process and activities represent only an idealization of the activities performed in practice. We refer to more detailed processes and activities (such as the process models we provide in this work).
>
>Detailed processes and activities only make sense within specific research areas or even families of experiments. For example, Arcuri \& Briand \cite{Arcuri-2014-randomized} have published recommendations for the statistical analysis of randomized algorithms. Arcuri \& Briand do not claim that their recommendations have general applicability. Other areas of SE, such as programming or testing, will need specific recommendations. Although this may seem strange or unusual, a cursory examination of experimental literature in other sciences demonstrates the opposite. For example, in Psychology, there are recommendations on how to measure motivation \cite{toure2014measure} or consider experience \cite{hoffman1998can}. In Medicine, there are models of pain \cite{petersen2002pain}. Other examples could be cited, but all of them agree on highlighting the specificities of a research area instead of hiding details in abstractions that ultimately prove inoperative.
>
>Finally, we must add specific competencies to processes, activities, and roles, such as negotiation skills, asset curation, or human resource management. Again, SE is lagging behind established trends in other areas. To cite a very relevant example, the European Union \cite{ResearchComp} has defined a framework containing seven competency areas, 38 competencies, and 389 learning outcomes for researchers. Other institutions have undertaken similar efforts.

We have also separated in specific sections the research *limitations* and the *future work*.

# Other issues

We have updated the *Acknowledgement* section to comply with the requirements imposed by our funding agency. The text of this section now reads:

> This work was supported by the grant PID2022-137846NB-I00 funded by MCIN/AEI/10.13039/501100011033, and by ''ERDF A way of making Europe'', and the Universidad de las Fuerzas Armadas - ESPE's Grupo de Investigación en Modelos de Producción de Software e Industria Inteligente (GrIMPSoftII).

### 2024-02-18 Article.tex
changes for V3

\usepackage[draft]{changes}

@@

\section*{Acknowledgment}
This work was supported by the project \replaced{PGC2018-097265-B-I00}{PID2022-137846NB-I00} funded by the Spanish Ministry of Science and Innovation and co-financed by FEDER, and the Universidad de las Fuerzas Armadas  - ESPE's Grupo de Investigación en Modelos de Producción de Software e Industria Inteligente (GrIMPSoftII). This work's preprint has been published \cite{Fonseca-2023-experimentation-proccess-SE-unpublished}.

>>>>

\usepackage[draft]{changes}
\definechangesauthor[name=version2, color=blue]{v2}
\definechangesauthor[name=version3, color=red]{v3}

@@

\section*{Acknowledgment}
This work was supported by the \deleted[id=v3]{project} \replaced[id=v3]{grant PID2022-137846NB-I00 funded by MCIN/AEI/10.13039/501100011033, and by ''ERDF A way of making Europe''}{PGC2018-097265-B-I00 funded by the Spanish Ministry of Science and Innovation and co-financed by FEDER}, and the Universidad de las Fuerzas Armadas - ESPE's Grupo de Investigación en Modelos de Producción de Software e Industria Inteligente (GrIMPSoftII). This work's preprint has been published \cite{Fonseca-2023-experimentation-proccess-SE-unpublished}.

### 2024-02-18 tex/conclusion.tex
changes for V3

\added{However, this work is subject to various limitations. The first and most important is the empirical basis obtained. Although ethnographic research has been extensive over time and conducted using multiple techniques, in the end, we only have data from one research group. Furthermore, this group's experimenters have self-trained in the field, meaning they have not received specific academic training. The observed peculiarities could be habits acquired over time that other, more recent groups with formal training do not possess.}

\added{We also cannot rule out errors in perception on the researchers' part as well. As mentioned earlier, we had yet to learn of experimentation at the beginning of ethnographic research, which may have led us to misinterpret some behaviors of the experimenters. On the other hand, perhaps the researchers' "fresh" perspective has allowed us to identify our findings.}

\deleted{We believe that the ''lab mythology'' is worth studying. If the differential characteristics of the theoretical vs. actual experiment process are specific to the research group, the inquiry will simply end. However, it is also possible that widespread patterns of behavior show up. Bringing such knowledge to the foreground could contribute to improving experimental practice.}

\deleted{We are aware that the findings of this study correspond to the particular perspective of an experimental research group, albeit representative of the ESE community. Therefore,}
\added{To overcome the limitations of the present investigation,} it is necessary to carry out further studies to generalize the results of the current work. \added{Our plans are as follows:}
\begin{itemize}

\item \added{Verify the existence of terminological diversity, one of the first aspects that has emerged in the research and is relatively easy to check. We have already started this work. We surveyed SE experimenters from multiple research groups. In the survey, we asked them about the meaning and usage of experimentation-related terms. Preliminary results confirm the existence of terminological diversity.}

\item \added{After completing the ethnographic research, we initiated a comparative study of the experimental process between SE and an established engineering discipline (specifically, biotechnology). Preliminary results indicate the absence of terminological diversity and a greater formalization of the experimental process within specific research areas but not at a general level. This could be the key to terminological diversity in SE: sub-areas of SE generally have not developed specific terminology and protocols (with \cite{Arcuri-2014-randomized} being a notable exception), which could lead to the need for ad-hoc concepts and processes.}

\item \replaced{Both works will likely be published shortly. In addition, we plan}{We plan} to conduct more studies \replaced{in the future}{(probably not ethnographic because they require too much time)} on this topic soon. \added{We still need to decide on the objects under study and the methodologies used since ethnography is very demanding. A possibility to study the impact of specialization in the sciences is to analyze experimental repositories. This represents an indirect but efficient approach in terms of time and effort. Another alternative is conducting qualitative studies using focus groups composed of SE experimental researchers.}

@@

\replaced{In short, w}{W}e believe that the "lab mythology" is worth studying. If the differential characteristics of the theoretical vs. actual experiment process \replaced{were}{are} specific to the research group, the inquiry \replaced{would}{will } simply end. However, it is also possible that widespread patterns of behavior show up. Bringing such knowledge to the foreground could contribute to improving experimental practice.

>>>>

\subsection{Implications for practice}

\added[id=v3]{Our findings suggest specific improvements that could be applied to SE experimentation. The first and most evident is transitioning from an experimental process primarily based on the application of experimental designs and the conduct of statistical analyses to a more professional approach based on explicit processes, activities, and roles, as is customary in many areas, e.g., Information Technology \cite{iso20000}.}

\added[id=v3]{It could be argued that current SE experimentation already has a well-defined process and activities, as can be seen, e.g., in the works of Juristo \& Moreno \cite{Juristo-2001-SE-experimentation} or Wohlin \textit{et al.} \cite{Wohlin-2000-Experimentation-SE}. However, as we have argued before, such a process and activities represent only an idealization of the activities performed in practice. We refer to more detailed processes and activities (such as the process models we provide in this work).}

\added[id=v3]{Detailed processes and activities only make sense within specific research areas or even families of experiments. For example, Arcuri \& Briand \cite{Arcuri-2014-randomized} have published recommendations for the statistical analysis of randomized algorithms. Arcuri \& Briand do not claim that their recommendations have general applicability. Other areas of SE, such as programming or testing, will need specific recommendations. Although this may seem strange or unusual, a cursory examination of experimental literature in other sciences demonstrates the opposite. For example, in Psychology, there are recommendations on how to measure motivation \cite{toure2014measure} or consider experience \cite{hoffman1998can}. In Medicine, there are models of pain \cite{petersen2002pain}. Other examples could be cited, but all of them agree on highlighting the specificities of a research area instead of hiding details in abstractions that ultimately prove inoperative.}

\added[id=v3]{Finally, we must add specific competencies to processes, activities, and roles, such as negotiation skills, asset curation, or human resource management. Again, SE is lagging behind established trends in other areas. To cite a very relevant example, the European Union \cite{ResearchComp} has defined a framework containing seven competency areas, 38 competencies, and 389 learning outcomes for researchers. Other institutions have undertaken similar efforts.}

\subsection{Limitations}

\added[id=v2]{However, this work is subject to various limitations. The first and most important is the empirical basis obtained. Although ethnographic research has been extensive over time and conducted using multiple techniques, in the end, we only have data from one research group. Furthermore, this group's experimenters have self-trained in the field, meaning they have not received specific academic training. The observed peculiarities could be habits acquired over time that other, more recent groups with formal training do not possess.}

\added[id=v2]{We also cannot rule out errors in perception on the researchers' part as well. As mentioned earlier, we had yet to learn of experimentation at the beginning of ethnographic research, which may have led us to misinterpret some behaviors of the experimenters. On the other hand, perhaps the researchers' "fresh" perspective has allowed us to identify our findings.}

\subsection{Future work}

\deleted[id=v2]{We believe that the ''lab mythology'' is worth studying. If the differential characteristics of the theoretical vs. actual experiment process are specific to the research group, the inquiry will simply end. However, it is also possible that widespread patterns of behavior show up. Bringing such knowledge to the foreground could contribute to improving experimental practice.}

\deleted[id=v2]{We are aware that the findings of this study correspond to the particular perspective of an experimental research group, albeit representative of the ESE community. Therefore,}
\added[id=v2]{To overcome the limitations of the present investigation,} it is necessary to carry out further studies to generalize the results of the current work. \added[id=v2]{Our plans are as follows:}
\begin{itemize}

\item \added[id=v2]{Verify the existence of terminological diversity, one of the first aspects that have emerged in the research and is relatively easy to check. We have already started this work. We surveyed SE experimenters from multiple research groups. In the survey, we asked them about the meaning and usage of experimentation-related terms. Preliminary results confirm the existence of terminological diversity.}

\item \added[id=v2]{After completing the ethnographic research, we initiated a comparative study of the experimental process between SE and an established engineering discipline (specifically, biotechnology). Preliminary results indicate the absence of terminological diversity and a greater formalization of the experimental process within specific research areas but not at a general level. This could be the key to terminological diversity in SE: sub-areas of SE generally have not developed specific terminology and protocols which could lead to the need for ad-hoc concepts and processes.}

@@

\replaced[id=v2]{Both works will likely be published shortly. In addition, we plan}{We plan} to conduct more studies \replaced[id=v2]{in the future}{(probably not ethnographic because they require too much time)} on this topic soon. \added[id=v2]{We still need to decide on the objects under study and the methodologies used since ethnography is very demanding. A possibility to study the impact of specialization in the sciences is to analyze experimental repositories. This represents an indirect but efficient approach in terms of time and effort. Another alternative is conducting qualitative studies using focus groups composed of SE experimental researchers.}

### 2024-02-18 tex/discussion.tex
changes for V3

	\item The ethnographic study revealed one consequence of such diversity. It is possible, albeit unlikely, that one experimenter performs all experimental activities. These activities are not only complex: they demand time. In a project (research or not), the diversity of activities is usually associated with roles with well-defined responsibilities. \textbf{Also in experimentation. We identified three roles}: (1) \textit{Research Manager}, (2) \textit{Experiment Manager}, and (3) \textit{Senior Experimenter}. These roles are probably related to the RGUS' size; smaller (or larger) groups may have more generic/specialized roles. The concept of~\textquotedblleft role\textquotedblright~applied to experimentation is unusual but not entirely new. Mohamed et al. \cite{Mohamed-1993-roles-ESE} proposed a framework to conduct ESE experiments. This framework was based on a merge between statistical and SE process concepts, \textbf{including explicit roles}.

@@

\item The replication problem is related to knowledge transfer \cite{Shull-2004-Knowledge-sharing-issues-SE}. Nevertheless, we wonder what exactly does it mean to transfer knowledge? To what extent or how should it be done? For example, Ferreira et al. \cite{Ferreira-2017-planning-experiments} emphasize methodological consistency over discussing methods and techniques.

>>>>

	\item The ethnographic study revealed one consequence of such diversity. It is possible, albeit unlikely, that one experimenter performs all experimental activities. These activities are not only complex: they demand time. In a project (research or not), the diversity of activities is usually associated with roles with well-defined responsibilities. \textbf{Also in experimentation. We identified three roles}: (1) \textit{Research Manager}, (2) \textit{Experiment Manager}, and (3) \textit{Senior Experimenter}. These roles are probably related to the RGUS' size; smaller (or larger) groups may have more generic/specialized roles. The concept of~\textquotedblleft role\textquotedblright~applied to experimentation is unusual but not entirely new. Mohamed \textit{et al.} \cite{Mohamed-1993-roles-ESE} proposed a framework to conduct ESE experiments. This framework was based on a merge between statistical and SE process concepts, \textbf{including explicit roles}.

@@

\item The replication problem is related to knowledge transfer \cite{Shull-2004-Knowledge-sharing-issues-SE}. Nevertheless, we wonder what exactly does it mean to transfer knowledge? To what extent or how should it be done? For example, Ferreira \textit{et al.} \cite{Ferreira-2017-planning-experiments} emphasize methodological consistency over discussing methods and techniques.

### 2024-02-18 tex/introduction.tex
changes for V3

\section{Introduction}\label{sec-introduction}
Replication is troublesome in general \cite{Klein-2018-many}, particularly in the social and life sciences \cite{Pashler-2012-perspectives} \cite{Baker-2016-lid-reproducibility}. In Empirical Software Engineering (ESE), the number of replications is small \cite{Bezerra-2015-Replication-SE-U-SMS} and the rate of confirmation of previous results is limited \cite{Jorgensen-2016-Incorrects-Results-SEE}.

Several reasons prevent replication. They have discussed elsewhere, e.g., \cite{Miller-2005-replicating-SE-experiments}, \cite{Demagalhaes-2015-replications-SE}. In this paper, we focus on a potential cause that, to our knowledge, has not been explored so far: How ESE researchers \textit{really} experiment.

@@

This paper aims to determine \textbf{how experimental researchers conduct experiments in practice}. We applied an ethnographic method \cite{Sharp-2016-Ethnographic-Studies-ESE} within an experienced experimental research group. We observed the researchers in their lab for two years. We created conceptual and process models to represent the concepts they use and the activities they perform daily. These models fit the community's procedures and terminology at a high level. Still, the models show particularities at lower levels: (1) The number and diversity of activities, (2) the existence of different roles, (3) the granularity of the concepts used by the roles, and (4) the viewpoints that different subareas or families of experiments have about the overall process.

>>>>

\section{Introduction}\label{sec-introduction}
Replication is troublesome in general \cite{Klein-2018-many}, particularly in the social and life sciences \cite{Pashler-2012-perspectives} \cite{Baker-2016-lid-reproducibility}. In Empirical Software Engineering (ESE)\added[id=v3]{, there is a consensus that replications are important \cite{shepperd2018role}, but} the number of replications is small \cite{Bezerra-2015-Replication-SE-U-SMS} and the rate of confirmation of previous results is limited \cite{Jorgensen-2016-Incorrects-Results-SEE}.

Several reasons prevent replication. They have discussed elsewhere, e.g., \cite{Miller-2005-replicating-SE-experiments} \cite{Demagalhaes-2015-replications-SE} \added[id=v3]{\cite{cockburn2020threats}}. In this paper, we focus on a potential cause that, to our knowledge, has not been explored so far: How ESE researchers \textit{really} experiment.

@@

This paper aims to \replaced[id=v2]{investigate}{determine} \textbf{how experimental researchers conduct experiments in practice}. \added[id=v2]{We intend to find out how experimental researchers plan, execute, analyze, and report their investigations and what concepts, protocols, and processes they use.}

We applied an ethnographic method \cite{Sharp-2016-Ethnographic-Studies-ESE} within an experienced experimental research group. We observed the researchers in their lab for two years. We created conceptual and process models to represent the concepts they use and the activities they perform daily. These models fit the community's procedures and terminology at a high level. Still, the models show particularities at lower levels: (1) The number and diversity of activities, (2) the existence of different roles, (3) the granularity of the concepts used by the roles, and (4) the viewpoints that different subareas or families of experiments have about the overall process.

### 2024-02-18 tex/research-execution.tex
changes for V3

\subsubsection{Analysis of the relevant literature}
The experimenters use the books by Wohlin et al. \cite{Wohlin-2000-Experimentation-SE} and Juristo \& Moreno \cite{Juristo-2001-SE-experimentation} as primary sources for their experimentation process. They also utilize specific papers that advise how to experiment in SE, e.g., Basili et al. \cite{Basili-1986-ESE}, and Kitchenham et al. \cite{Kitchenham-2002-empirical-research-guidelines-SE}. Basili et al.'s work \cite{Basili-1999-families-experiments} was also considered relevant due to the experimenters' interest in carrying out experimental replications.

@@

\begin{itemize}
	\item Wohlin et al. \cite{Wohlin-2000-Experimentation-SE}: "Experiment idea, Experiment scoping, Experiment planning, Experiment operation, Analysis \& interpretation, Presentation \& package, and Experiment report."
	\item Juristo \& Moreno \cite{Juristo-2001-SE-experimentation}: "Objective Definition, Design, Execution, and Analysis."
	\item Basili et al. \cite{Basili-1986-ESE}: "Definition, Planning, Operation, and Interpretation."
	\item Kitchenham et al. \cite{Kitchenham-2002-empirical-research-guidelines-SE}: "Experimental context, Experimental design, Conduct of the experiment and Data Collection, Analysis, Presentation of results, and Interpretation of results."
\end{itemize}

We looked for additional ESE books to verify whether terminological diversity was a widespread problem. We only found a new edition of Wohlin et al.'s book \cite {Wohlin-2012-experimentatio-SE}, and the specialized books by Runeson et al. \cite{Runenson-2012-case-study-SE} and Kitchenham et al. \cite{Kitchenham-2015-Evidence-Based-SE}. As expected, these books share standard terms but use diverse terms throughout the text, contributing to the terminological diversity.

@@

\begin{figure}[htbp!]
	\centering
	\includegraphics[width=\columnwidth]{images/Process-Approximation-Draft}
	\caption{Example of a sketch made by the experimenters. \added{The researchers have used drafts like this to progress in the definition of the experimental process workflow described in Section~\ref{sec-results:process-workflow}.}}
	\label{fig-proceso-exp-boseto}
	\highlight{This new figure replaces the previous Figure 2, which was included}
	\highlight{in the first version of the manuscript.}
\end{figure}

>>>>

\subsubsection{Analysis of the relevant literature}
The experimenters use the books by Wohlin \textit{et al.} \cite{Wohlin-2000-Experimentation-SE} and Juristo \& Moreno \cite{Juristo-2001-SE-experimentation} as primary sources for their experimentation process. They also utilize specific papers that advise how to experiment in SE, e.g., Basili \textit{et al.} \cite{Basili-1986-ESE}, and Kitchenham \textit{et al.} \cite{Kitchenham-2002-empirical-research-guidelines-SE}. Basili \textit{et al.}'s work \cite{Basili-1999-families-experiments} was also considered relevant due to the experimenters' interest in carrying out experimental replications.

@@

\begin{itemize}
	\item Wohlin \textit{et al.} \cite{Wohlin-2000-Experimentation-SE}: "Experiment idea, Experiment scoping, Experiment planning, Experiment operation, Analysis \& interpretation, Presentation \& package, and Experiment report."
	\item Juristo \& Moreno \cite{Juristo-2001-SE-experimentation}: "Objective Definition, Design, Execution, and Analysis."
	\item Basili \textit{et al.} \cite{Basili-1986-ESE}: "Definition, Planning, Operation, and Interpretation."
	\item Kitchenham \textit{et al.} \cite{Kitchenham-2002-empirical-research-guidelines-SE}: "Experimental context, Experimental design, Conduct of the experiment and Data Collection, Analysis, Presentation of results, and Interpretation of results."
\end{itemize}

We looked for additional ESE books to verify whether terminological diversity was a widespread problem. We only found a new edition of Wohlin \textit{et al.}'s book \cite {Wohlin-2012-experimentatio-SE}, and the specialized books by Runeson \textit{et al.} \cite{Runenson-2012-case-study-SE} and Kitchenham \textit{et al.} \cite{Kitchenham-2015-Evidence-Based-SE}. As expected, these books share standard terms but use diverse terms throughout the text, contributing to the terminological diversity.

@@

\begin{figure}[htbp!]
	\centering
	\includegraphics[width=\columnwidth]{images/Process-Approximation-Draft}
	\caption{Example of a sketch made by the experimenters. \added[id=v2]{The researchers have used drafts like this to progress in the definition of the experimental process workflow described in Section~\ref{sec-results:process-workflow}.}}
	\label{fig-proceso-exp-boseto}
	\highlight{This new figure replaces the previous Figure 2, which was included}
	\highlight{in the first version of the manuscript.}
\end{figure}

### 2024-02-18 tex/research-method.tex
changes for V3

\section{Research Method}\label{sec-research-method}
Ethnography is a \textit{participant observation} method. It does not interfere (beyond the ethnographic researchers' presence) with the observed ESE researchers' "life" \cite{Sharp-2016-Ethnographic-Studies-ESE}. Ethnographic researchers adopt "the ESE researchers' point of view" to better understand the reason for their actions \cite[pp. 55-56]{Outhwaite-2007-sage}. Thus, the biases and misperceptions that ESE researchers might have about themselves can be detected. Ethnography also captures the rich context that surrounds ESE researchers.

Ethnography identifies details that are ignored by or in other methods. The disadvantage is the time required for learning \cite{Sharp-2016-Ethnographic-Studies-ESE}. Thus, we opted to study a single experimental research group over a long period instead of working halfway with several experimental research groups over shorter periods. This decision was also motivated by the challenge of finding several open-minded experimental research groups willing to participate in this research. We selected a representative, i.e., respected and well-known, research group to guarantee (to the extent possible) the reliability of the results, their validity, and the findings' scientific significance \cite{sjoberg-2007-future-empical-methods}.

We followed Runenson et al.'s case study guidelines \cite{Runenson-2012-case-study-SE} wherever they apply to ethnographical studies. Our research protocol is organized along three axes:

>>>>

\section{Research Method}\label{sec-research-method}

Ethnography is a \textit{participant observation} method \added[id=v2]{used with some frequency in SE, e.g., \cite{traini2022exploring, azarova2022just}}. It does not interfere (beyond the ethnographic researchers' presence) with the observed ESE researchers' "life" \cite{Sharp-2016-Ethnographic-Studies-ESE}. Ethnographic researchers adopt "the ESE researchers' point of view" to better understand the reason for their actions \cite[pp. 55-56]{Outhwaite-2007-sage}. Thus, the biases and misperceptions that ESE researchers might have about themselves can be detected. Ethnography also captures the rich context that surrounds ESE researchers.

Ethnography identifies details that are ignored by \deleted[id=v2]{or in} other method\replaced[id=v2]{ologies, including Artificial Intelligence (AI) \cite{yadav2023development}}{s}. The disadvantage is the time required for learning \cite{Sharp-2016-Ethnographic-Studies-ESE}. Thus, we opted to study a single experimental research group over a long period instead of working halfway with several experimental research groups over shorter periods. This decision was also motivated by the challenge of finding several open-minded experimental research groups willing to participate in this research. We selected a representative, i.e., respected and well-known, research group to guarantee (to the extent possible) the reliability of the results, their validity, and the findings' scientific significance \cite{sjoberg-2007-future-empical-methods}.

We followed Runenson \textit{et al.}'s case study guidelines \cite{Runenson-2012-case-study-SE} wherever they apply to ethnographical studies. Our research protocol is organized along three axes:

### 2024-02-18 tex/threats-to-validity.tex
changes for V3

\section{Threats to Validity}\label{sec-threats}
Validity is the degree to which the research is conducted transparently and without biases. We have followed Runeson et al.'s recommendations \cite[p. 71–73]{Runenson-2012-case-study-SE} to design the ethnographic research and describe the threats to validity. Runeson et al., following Yin \cite{Yin-2009-case-study}, classify threats into four groups: (1) construct validity, (2) internal validity, (3) external validity, and (4) reliability.

>>>>

\section{Threats to Validity}\label{sec-threats}
Validity is the degree to which the research is conducted transparently and without biases. We have followed Runeson \textit{et al.}'s recommendations \cite[p. 71–73]{Runenson-2012-case-study-SE} to design the ethnographic research and describe the threats to validity. Runeson \textit{et al.}, following Yin \cite{Yin-2009-case-study}, classify threats into four groups: (1) construct validity, (2) internal validity, (3) external validity, and (4) reliability.
